---
title: "Stepwise2"
author: "Zelin Wang"
date: "2023-09-09"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load data
```{r}
cs_data<-read.csv("CS.csv")
```

# Set random seed for subsequent random selection and assignment operations
```{r}
set.seed(2023)
```

# Partition data and create index matrix of selected values
```{r}
library(caret)
index <- createDataPartition(cs_data$os, p=.8, list=FALSE, times=1)
```

# Create test and train data frames
```{r}
train_df <- cs_data[index,]
test_df <- cs_data[-index,]
```

# Verify number of rows (cases) in each data frame
```{r}
nrow(train_df)
nrow(test_df)
```

#define response variable
```{r}
y_nom <- train_df$os
```

#define matrix of predictor variables
```{r}
x_nom <- data.matrix(train_df[, c(2:21)])
data_nom <- data.frame(x_nom,y_nom)
```

# Generate full model, exclude all cod related variables, and numeric variables.
```{r}
full_model<-glm(y_nom~as.factor(gender)+as.factor(race)+as.factor(age)+as.factor(size)+as.factor(marry)+as.factor(income)+as.factor(site)+as.factor(grade)+as.factor(kind)+as.factor(T)+as.factor(N)+as.factor(surgery_pri)+as.factor(RX_Summ)+as.factor(radiate)+as.factor(chem)+as.factor(CEA)+as.factor(bone)+as.factor(brain)+as.factor(lung)+as.factor(group),data=data_nom,family = binomial(link="logit"))
summary(full_model)
```

# Perform stepwise regression
## create null model
```{r}
null_model<-glm(y_nom~1,data=data_nom,family=binomial(link="logit"))
```

## Perform forward selection
### set seed
```{r}
set.seed(2023)
```

### AIC threshold (Akaike Information Criterion)
```{r}
model_step1 = step(null_model, 
                  scope = list(lower=null_model, upper=full_model),
                  direction = "forward", k= 2, trace=0) 
```

```{r}
###summary model
summary(model_step1)
cat("AIC = ", AIC(model_step1), sep = "")
```

### BIC threshold (Bayesian information criterion)
```{r}
model_step2 = step(null_model, 
                  scope = list(lower=null_model, upper=full_model),
                  direction = "forward", k= log(nrow(cs_data)), trace=0) 
```
```{r}
#summary model
summary(model_step2)
cat("BIC = ", BIC(model_step2), sep = "")
```


## Perform backward elimnation
### AIC threshold (Akaike Information Criterion)
```{r}
model_step3 = step(full_model, 
                  direction = "backward", k= 2, trace=0) 
```

```{r}
#summary model
summary(model_step3)
cat("AIC = ", AIC(model_step3), sep = "")
```

### BIC threshold (Bayesian information criterion)
```{r}
model_step4 = step(full_model, 
                  direction = "backward", k= log(nrow(cs_data)), trace=0) 
```

```{r}
#summary model
summary(model_step4)
cat("bIC = ", BIC(model_step3), sep = "")
```


## Perform stepwise selection (both)
### AIC threshold (Akaike Information Criterion)
```{r}
model_step5 = step(null_model, 
                  scope = list(lower=null_model, upper=full_model),
                  direction = "both", k= 2, trace=0) 
```

```{r}
#summary model
summary(model_step5)
cat("AIC = ", AIC(model_step5), sep = "")
```

### BIC threshold (Bayesian information criterion)
```{r}
model_step6 = step(null_model, 
                  scope = list(lower=null_model, upper=full_model),
                  direction = "both", k= log(nrow(cs_data)), trace=0)
```

```{r}
#summary model
summary(model_step6)
cat("BIC = ", BIC(model_step6), sep = "")
```








